---
title: "hw03"
author: "Ziqiang Tang"
data: "2018-09-30"
output: github_document
---

## Overview
The aim of this aissignment is to manipulate and explore a dataset with the dplyr package, complmented by visulization made with ggplot2.


## Task menu
First, let's library all dataset.
```{r}
library(gapminder)
library(tidyverse)
library(knitr)
```

###1. Get the maximum and minimum of GDP per capital for all continents
```{r}
GDP_per_cap <- gapminder %>% 
  group_by(continent) %>% 
  summarize(minGDP = min(gdpPercap),
            maxGDP = max(gdpPercap))
GDP_per_cap%>% 
    kable(col.names = c("Continent", "Min GDP", "Max GDP"))
```
#### Let's try to view the minimum and maximum of GDP per capital for all continents.
```{r}
GDP_per_cap %>% 
  ##get the col 2 and 3 data_set
  gather(key=type, value=GDP_per_cap, 2:3) %>%  
  ##plot the data with color fill
  ggplot(aes(x = continent, y = GDP_per_cap, fill = continent)) +
  ##the height of bars reporesents the data
  geom_col(alpha=0.5)+
  ##warp with maxGDP and minGDP with the y-scale is not aligned
  facet_wrap(.~type)  +
  ggtitle("The GDP per capital for all continents") +
  xlab("Continent") +
  ylab("GDP per capital")
```
####Let's try to view the average of GDP per capital for all continents.
```{r}
GDP_per_cap %>% 
ggplot(aes(x = minGDP, y = maxGDP)) +
    scale_size_area() +
  geom_point(aes(color=continent, size = (maxGDP - minGDP)), 
             alpha =0.5) +
  coord_polar() +
  ggtitle("maxGDP - minGDP per capital in all continents")
```
This polar diagram give us the intuition of the largeset GDP difference continent is Oceania while the smallest GDP difference is Africa.

###2. Look at the spread of GDP per capital within the continents
